FACE VALUE: THE POWER OF IMAGES AT APHRODISIAS. How Digital Resources Can Transform our use of Palaeography in understanding Inscriptions.

### 2015-12-16 source/10/Paper proposal for EAGLE final.tex
started work on tex of papers, added folders for CNR papers, changed main stracture to include keynotes and AHE abstracts

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{}
\author{Abigail S Graham}
\date{2015-11-16}
\begin{document}
\begin{enumerate}
\item FACE VALUE: THE POWER OF IMAGES AT APHRODISIAS
\end{enumerate}
How Digital Resources Can Tranform our use of Palaeography in understanding Inscriptions.

\bigskip

\ Dr. Abigail Graham

\bigskip

The University of Warwick (Lecturer)

\bigskip

Abigail.graham@warwick.ac.uk

\bigskip

Possibly for subject panel: Dating criteria for epigraphic material: possible approaches.

\bigskip

Abstract:

\bigskip

\ This paper will consider how the publication of a large digital corpus (The Inscriptions of Aphrodisias [Reynolds,
Roueche \& Bodard, 2007] has shaped the assessment of inscriptions, particularly regarding palaeography and the dating
of inscriptions. A case study of dedicatory inscriptions from the Temple of Aphrodite at Aphrodisias will explore how
our approach to palaeography and dating has evolved with digital resources, identifying areas where challenges remain
and considering how improvements could be made in both our approach to and in publication of epigraphic materials.

\bigskip

Keywords

\bigskip

Dating, Palaeography, Ordinatio, Letter forms, Aphrodisias, Context, Recarving.

\bigskip

\bigskip

1.1 Introduction

\bigskip

\begin{enumerate}
\item \begin{enumerate}
\item \begin{enumerate}
\item Caveat Lector: \ Defining Palaeography and traditional approaches
\end{enumerate}
\end{enumerate}
\end{enumerate}

\bigskip

As the co-ordinator and lecturer on three graduate level courses of Roman Epigraphy, I am invariably asked the same
question: how do you use letterforms to date inscriptions? My answer is always the same: “very carefully”. \ Studying
palaeography within the discipline of ancient epigraphy can be a journey into thorny hedge where one can easily “fall
into that category of human endeavor known as stylistic attribution and inevitably involve subjectivity”.\footnote{
Tracy, 1995, 3-4. } Stephen Tracy advises that palaeographic surveys should be carried out with caution: “Caveat Lector
must needs be our motto”. \ The study of palaeography is problematic on a number of levels, both in the way it is
defined and the ways in which it is employed. Many scholars dismiss lettering as a means of dating and they make an
important point: dating by a single criterion, especially a stylistic one, is somewhat precarious. When studied in
isolation, letterforms present stylistic variations that may be characteristic of a specific individual, workshop or an
urban area. While observations about carving techniques can be helpful in specific case studies, they are more
problematic when applied generally on a broader scale (e.g. to larger geographical areas or time trame) where
archaisms, local styles, and variations can create distortions.\footnote{ Susini 1973; Petrucci 1993; Di Stefano
Manzella 1995 163-181, Cooley 2012, 433; Edmundson 2015 122-125.} Similar caution would be applied in dating a
sculpture on the basis of a hair fragment. Analysis of statues considers a number of factors, material, hairstyle,
drapery and/or context. Inscriptions are doubly difficult, as they fall into categories of both text and an object.
While there is potentially more information, there is also a greater chance that it will be contradictory. Thus the use
of lettering, an “imprecise science”, is better used in combination with a number of different factors.\footnote{
Harris 1989, 26-9; Bodel 2001 3-5; Di Stefano Manzella 2007; 393-418 and Cooley 2012, 432-435. } Despite the
aforementioned limitations, a number of informative palaeographic studies have been produced. The success of these is
based on detailed commentaries on a specific corpus of material, a transparent methodology and the incorporation of
numerous high quality images.\footnote{ For example: Gordon 1957 offers more than 50 plates and figures, Di Stefano
Manzella 1987 provides 218 illustrations, Tracy 1990, 1995 and 2003 uses over 60 plates and figures in each volume. \ }

\bigskip

A second issue in the study of palaeography involves the access and publication of epigraphic materials. \ Access to
large corpora of inscriptions has traditionally been limited to a small audience of scholars and site visitors. Apart
from a few museum collections, which happen to have inscriptions arranged in roughly chronological order,\footnote{
Perhaps the finest example of a museum in which one can gain and understanding of palaeography (and inscriptions as
whole) is the Museo Epigrafico Nazionale, Roma. } there are few places where one can visually experience the
development of carving styles over time. For the lucky few who attain access (and permission) to study a large corpus
of inscriptions, publishing these texts with supporting images can be challenging and expensive. Studies of
palaeography in ancient inscriptions have often been, by necessity, selective with images making it difficult for both
the author and readers to develop a detailed understanding of carving trends and practices. In this traditional format,
inscriptions were also separated from their archaeological context and the accompanying artwork while the reader,
viewing only the lettering, was often removed from the visual elements of the inscription (e.g. the type of stone, use
of spacing and decoration, letter size). \ This is a suboptimal way of assessing epigraphic evidence.

\bigskip

The advent of online corpora have increased both the access and the development of discussions regarding epigraphic
monumentality, including new methodologies, approaches as well as attempts to redefine genre classifications and
terminology.\footnote{ Woolf 1996, 22-39, Eck 2010, Panciera 2012, 1-10, Graham 2013, 1-17. \ For a recent collection
and analysis of epigraphic databases, see Tom Elliot’s work \ (chapter 5) in Bruun and Edmundson 2015, 78-85. \ } \ A
number of recent studies have used changes in the appearance of inscriptions over time, such as the use of different
media, decorative and paratextual elements\footnote{ \ One marks in the text see Susini 1973, 26 and Cooley in Jansen
2014, 143-155.} (e.g. ivy leaves as interpuncts (hederae distinguentes),\footnote{ \ Hommel 1970, 293-303. }
abbreviations\footnote{ Gordon, “Supralineate Abbreviations” 1957, 111.} and spacing between words\footnote{ On using
space to reconstruct inscriptions see Alföldy 1995, 195-226, Grasby 2002, 151-176.} alongside lettering, as dating
criteria.\footnote{ For an extensive list of dating conventions cf..Di Stefano Manzella 1987, Chapter 20. }
\ Panciera’s recent article, in particular, advocates the significance of the public context and visibility of
inscriptions.\footnote{ “I would propose to regard as an ‘inscription’ any particular type of written human
communication of the sort that we would today call unidirectional…not being addressed to a person or to a group but to
a collectivity, and for this reason is made with the location, writing technique, graphic form and impagination, mode
and register of expression chosen because they are most suitable to the attainment of its intended goal.” (Panciera
2012: 8).} \ In this vein, it is worth considering how the physical characteristics of an inscription belong within a
broader assessment of a culture of writing. Is palaeography the study of letterforms alone? Was a focus on letterforms
a deliberate choice or a product of the traditional constraints in accessibility and publication of epigraphic
materials? The following case study from Aphrodisias will explore these questions further.

\bigskip

1.1.3 Outline

\bigskip

The focus of this paper is a series of rather unimpressive column dedications from the temple of Aphrodite at
Aphrodisias. \ While the information recorded in the texts is unremarkable, the journey of these inscriptions from
their publication as part of an epigraphic corpus to a recent digital publication reveals an epic transformation in the
format and approach to these materials. \ This study will begin with brief overview of previous published editions
(e.g. MAMA volume VIII, 1962) followed by an assessment of the information available on the current Inscriptions of
Aphrodisias (2007) website. Through careful assessment of text (including formula, vocabulary and spelling) and its
presentation (the arrangement of the text, use of decoration and spaces, as well as lettering) in the images provided,
this survey will demonstrate how the availability of published images and inclusion of dating criteria have increased
the amount of information available whilst also adding clarity to the process of dating an inscription. By examining
how we define and use palaeography to evaluate inscriptions on the Inscriptions of Aphrodisias website (IAph 2007), we
can observe how the discipline has evolved and what changes may be possible in approaches and the publication of
epigraphic material.

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

[Warning: Draw object ignored]

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\begin{enumerate}
\item \begin{enumerate}
\item Publishing inscriptions: A Case Study of Column Dedications from The Temple Of Aphrodite at Aphrodisias 
\end{enumerate}
\end{enumerate}

\bigskip

1.2.1 A brief introduction to the inscriptions IAph 1.4-1.6.

\bigskip

Three inscriptions, each of which record the dedication of a column at the Temple of Aphrodite, will be the subject of
this survey. \ The first two texts were noted as early as the 18th century and copied in a notebook by the British
Architect Deering in the early 19th century. A third version of the text, was \ uncovered during excavations at the
temple site by the French Engineer Gaudin in the early 20th century.\footnote{ A comprehensive history of the history
and bibliography, as well as a description of the resources can be found on the IAph website:
http://insaph.kcl.ac.uk/iaph2007/iAph010006.html. } The inscriptions are recorded on tabella ansata (0.745 x 0.465m) as
part of fluted marble columns of the peristasis, some of these have been reconstructed in modern restorations on the
site.

\bigskip

These texts can be dated through a number of different criteria. \ Contextual association with construction of the
temple, dates between the end of the 1st century BC and the early 1st century AD, though the land was clearly in use
well before this time. \footnote{ Smith 1996, 41-42 and Reynolds 1990, 37. References to Sulla’s dedications are in
Appian BC 1.97. \ Caesar and Augustus’ acknowledgements of the sacred space are evident in IAph 8.27, 8.31 and 1.1. }
\ This is corroborated by coins depicting the temple, which date from 2 BC- 14 AD.\footnote{ Coin type 41 depicts
Augustus (OBV) and the temple of Aphrodite (REV) in MacDonald 1992, Plate V R131. } Prosopography is also informative:
Gaius Julius Zoilos (freed by Caesar or Octavian), who dedicated the theatre at Aphrodisias ca. 28 BC, also dedicated
the naos of the temple (IAph 1.2), perhaps posthumously. \footnote{ The titles on the architrave inscription, soter and
euergetes, are absent from Zoilos’ other inscriptions and may imply a posthumous dedication, perhaps by the boule and
demos (Reynolds 1990, 38).} The benefactor of the columns in this case study, Eumachos Diogenes, also comes from an
established family that flourished well into the 2nd century AD.\footnote{18 The success of Eumachus Diogenes’ family,
which included the first known Aphrodisian to hold a procuratorship in the 2nd century AD, reveals an enduring
significance for his family monuments in the city (Reynolds 1999b, 327-334).} \ Finally, the formula of the text the
inscription and the vocabulary, particularly the term philokaisar, suggest a late Republican or Augustan
date.\footnote{ The formula of the dedication, which lists the benefactor first is evident throughout the Hellenistic
period in Aphrodisias (and Asia Minor) until the early Imperial Period, after which a new formula (beginning with
recipients (e.g. Aphrodite and Imperial recipients) is predominant (cf. Graham 2013, 4-7). \ The Augustan and early
Tiberian uses of the philokaisar are known from a dedication at Ioulis dated ca. 27 BC - AD 14 (SEG XLVIII (1998) no.
1129) and in a monument to Ti. Cl. Drusus in Patara (SEG XLIV (1994) no. 1205). For significance and date of the title
see Buraselis 2001, 101-105. } \ Before encountering the inscriptions face to face, the dating, function and meaning of
these inscriptions appears to be quite straightforward. \ So let us examine the experience of viewing these
inscriptions various published formats.

\bigskip

1.2.2. Publishing IAPH 1.4 and 1.5 in MAMA VIII (nos. 347 \& 348): A series of copies?

\bigskip

Cormack published the first comprehensive catalogue of inscriptions from Aphrodisias as part of his Monumenta Asiae
Minoris Antiqua volume VIII in 1962. \ This was a great undertaking that included texts as well as images. Two of the
three texts \ (IAph 1.4 \& 1.5) were published here as MAMA 437 and MAMA 438 respectively, while the third text was
merely mentioned as a further copy.

\bigskip

\bigskip

\bigskip

\bigskip

[Warning: Draw object ignored][Warning: Draw object ignored]

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

[Warning: Draw object ignored]

\bigskip

The two published texts reveal similar inscriptions with a few variations.\footnote{ MAMA 437 and 438 are represented in
their digital format as published the Packhum website \ (Aphrodisias 108 and 109 (respectively)).
http://epigraphy.packhum.org/inscriptions/main?url=oi\%3Fikey\%3D256987} \ IAph 1.4 (MAMA 437) has a slightly less
impressive ordinatio, particularly in line 4 with an odd line break, a misspelling of Ammias (as Amias) and Adrastus
(line 5) with a sigma omitted. It is difficult to determine how or why these errors occurred without indications of
spacing or an examination of the stone. \ One could argue that spelling and arrangement of the text were not important
to a broader audience at Aphrodisias, however, this is negated by a second copy of the dedication (IAph 1.5, MAMA 438),
which portrays a more skillful execution and arrangement of the text. Spelling errors are rectified and the arrangement
of the text creates a distinction in Line 5 between Eumachus and Ammias, and for the demos, which is isolated on the
last line. These subtle variations only emerge with a close reading of the text, and it would be easy to overlook them.
The published format of the texts also makes them appear similar in lettering and arrangement (e.g. left indentation).
\

\bigskip

@@

\bigskip

\ The third text version of this text is mentioned in the commentary and based on the information provided (two similar
texts and a reference to a “copy”) one would expect that it was very similar. The same editorial practice is employed
in the parallel column dedications from a different benefactor (I.Aph 7 \& I.Aph 8), where 2nd version of the text is
noted (IAph 1.8) but only one (I.Aph 1.7, MAMA 450) was published. The apparent verisimilitude of the column
dedications is corroborated by the published images, which depict two inscriptions (MAMA 437 and MAMA 450) with similar
late 1st c. BC/early 1st c. AD lettering. When publishing a large corpus of inscriptions, referring to copies is
understandable. However, treating texts as “copies” as opposed to individual monuments can, and in this case will,
prove problematic.\footnote{ For another study of “copy” inscriptions at Aphrodisias see Graham 2015 (forthcoming).}
While MAMA VIII offered a broader audience access to the inscriptions at Aphrodisias, the approach to the inscriptions
and the quality of the images also imposed limitations that made it difficult to analyse the physical elements of the
inscription or to question the proposed texts and restorations. \

1.2.3 \ Living in a Digital World: The Publication of Column Dedications in IAph 2007 (IAph 1.4.)

\bigskip

The Inscriptions of Aphrodisias project was a groundbreaking endeavour both for publishing epigraphic materials and
employing Epi Doc conventions with XML for marking up ancient documents. Its success has inspired a number of
significantly larger projects with international scope and collaboration. The approach to the materials is detailed
with categories covering physical characteristics, history of discovery and bibliography (which is important given the
wealth of superlative scholarship).\footnote{ It is worth noting that websites, however they develop, will remain
supplementary to published scholarship.} \ In addition to search functions, the page is more interactive, the texts
have been revisited, translations added, as well as commentary with weblinks to parallel texts. The organization of the
inscriptions by context allows the reader to view the inscriptions in context and to gain a better understanding of the
area’s “epigraphic habit” (Figure 1: Screenshot of \ IAph 4 webpage).

\bigskip

 \includegraphics[width=14.605cm,height=8.726cm]{PaperproposalforEAGLEfinal-img001.png}

\bigskip

\bigskip

@@

\bigskip

The published texts of IAph 1.4 and MAMA 437 present one significant variation: the inclusion of space indicators within
the text. This practice allows the reader to see how the use of space relates with the text of the inscription,
particularly in the case of line 4, where challenges in the text (misspelling of Ammias and carrying over of a single
letter from the previous line), can be observed in the inclusion of a vacat at the end of the line.\footnote{ Vacats,
more sparingly employed in late Republican and Augustan inscriptions at Aphrodisias, usually serve a grammatical and/or
decorative function (e.g. giving distinction to a name or key elements/individuals in the text (cf. building
dedications at the Sebasteion IAph 9.1, 9.25, 9.112). This vacat serves little grammatical function, and is likely to
be a result of the arrangement of the text/ omission in Ammias’ name. } \ The indentation on the last line also reveals
a left orientation of the line which is more common in Late Republican/Augustan texts at Aphrodisias (e.g. IAph 1.1,
1.7, 1.8 and 1.38), as opposed to the “justified” approach (indentations on both sides of the text to accentuate a name
or word), which is more common in later Imperial inscriptions at Aphrodisias.

\bigskip

\bigskip

\bigskip

[Warning: Draw object ignored]

[Warning: Draw object ignored]

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

\bigskip

High quality images on the website facilitate connections between the published text and the inscription. In the image
of IAph 1.4 (Figure 2. Photograph of IAph 1.4), one can to observe how descriptive elements manifest themselves in an
inscription: how varying letter sizes (2-3 cm) represent a lack of uniformity and crowding as space becomes more
cramped, and how, after the beautifully spaced upper lines (lines 1-3) the carver struggled (line 4) to fit the letters
in, possibly noticing his spelling error only when there was extra space at the end of the line. \

\bigskip

\bigskip

 \includegraphics[width=8.758cm,height=7.668cm]{PaperproposalforEAGLEfinal-img002.jpg}

\bigskip

>>>>

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{FACE VALUE: THE POWER OF IMAGES AT APHRODISIAS. How Digital Resources Can Transform our use of Palaeography in understanding Inscriptions.}
\author[war]{Abigail S. Graham\corref{first}}
\address[war]{The University of Warwick}
\cortext[first]{Corresponding author. Email: Abigail.graham@warwick.ac.uk}
\date{2015-11-16}
\begin{document}
 \maketitle
\begin{abstract}
This paper will consider how the publication of a large digital corpus The Inscriptions of Aphrodisias \citet{ReynoldsRouecheBodard2007} has shaped the assessment of inscriptions, particularly regarding palaeography and the dating
of inscriptions. A case study of dedicatory inscriptions from the Temple of Aphrodite at Aphrodisias will explore how
our approach to palaeography and dating has evolved with digital resources, identifying areas where challenges remain
and considering how improvements could be made in both our approach to and in publication of epigraphic materials.

\end{abstract}
\keywords{Dating, Palaeography, Ordinatio, Letter forms, Aphrodisias, Context, Recarving.}

\section{Introduction}
\subsection{Caveat Lector: Defining Palaeography and traditional approaches}

\noindent  As the co-ordinator and lecturer on three graduate level courses of Roman Epigraphy, I am invariably asked the same
question: how do you use letterforms to date inscriptions? My answer is always the same: ``very carefully''. Studying
palaeography within the discipline of ancient epigraphy can be a journey into thorny hedge where one can easily ``fall
into that category of human endeavor known as stylistic attribution and inevitably involve subjectivity''.\footnote{\citet[3-4]{tracy1995}} Stephen Tracy advises that palaeographic surveys should be carried out with caution: ``Caveat Lector
must needs be our motto''. The study of palaeography is problematic on a number of levels, both in the way it is
defined and the ways in which it is employed. Many scholars dismiss lettering as a means of dating and they make an
important point: dating by a single criterion, especially a stylistic one, is somewhat precarious. When studied in
isolation, letterforms present stylistic variations that may be characteristic of a specific individual, workshop or an
urban area. While observations about carving techniques can be helpful in specific case studies, they are more
problematic when applied generally on a broader scale (e.g. to larger geographical areas or time trame) where
archaisms, local styles, and variations can create distortions.\footnote{\citet{susini1973}; \citet{petrucci1993}; \citet[163-181]{Manzella1995}; \citet[433]{cooley2012}; \citet[122-125]{bruun2015}.} Similar caution would be applied in dating a
sculpture on the basis of a hair fragment. Analysis of statues considers a number of factors, material, hairstyle,
drapery and/or context. Inscriptions are doubly difficult, as they fall into categories of both text and an object.
While there is potentially more information, there is also a greater chance that it will be contradictory. Thus the use
of lettering, an ``imprecise science'', is better used in combination with a number of different factors.\footnote{\citet[26-9]{harris1989}; \citet[3-5]{bodel2001}; \citet[393-418]{Manzella2007} and \citet[432-435]{cooley2012}.} Despite the
aforementioned limitations, a number of informative palaeographic studies have been produced. The success of these is
based on detailed commentaries on a specific corpus of material, a transparent methodology and the incorporation of
numerous high quality images.\footnote{ For example: \citet{Gordon1957} offers more than 50 plates and figures, \citet{Manzella1987} provides 218 illustrations, \citet{tracy1990,tracy1995,tracy2003} uses over 60 plates and figures in each volume.}

A second issue in the study of palaeography involves the access and publication of epigraphic materials. Access to
large corpora of inscriptions has traditionally been limited to a small audience of scholars and site visitors. Apart
from a few museum collections, which happen to have inscriptions arranged in roughly chronological order,\footnote{
Perhaps the finest example of a museum in which one can gain and understanding of palaeography (and inscriptions as
whole) is the Museo Epigrafico Nazionale, Roma. } there are few places where one can visually experience the
development of carving styles over time. For the lucky few who attain access (and permission) to study a large corpus
of inscriptions, publishing these texts with supporting images can be challenging and expensive. Studies of
palaeography in ancient inscriptions have often been, by necessity, selective with images making it difficult for both
the author and readers to develop a detailed understanding of carving trends and practices. In this traditional format,
inscriptions were also separated from their archaeological context and the accompanying artwork while the reader,
viewing only the lettering, was often removed from the visual elements of the inscription (e.g. the type of stone, use
of spacing and decoration, letter size). This is a suboptimal way of assessing epigraphic evidence.

The advent of online corpora have increased both the access and the development of discussions regarding epigraphic
monumentality, including new methodologies, approaches as well as attempts to redefine genre classifications and
terminology.\footnote{\citet[22-39]{woolf1996}, \citet{eck2010}, \citet[1-10]{panciera2012}, \citet[1-17]{graham2013}. For a recent collection and analysis of epigraphic databases, see \citet[78-85]{Elliot2015}.} A
number of recent studies have used changes in the appearance of inscriptions over time, such as the use of different
media, decorative and paratextual elements\footnote{On marks in the text see \citet[26]{susini1973} and \citet[143-155]{cooley2014}.} (e.g. ivy leaves as interpuncts (hederae distinguentes),\footnote{\citep[293-303]{hommel1970}}
abbreviations\footnote{\citep[111]{Gordon1957}} and spacing between words\footnote{On using
space to reconstruct inscriptions see \citet[195-226]{alfoldy1995}, \citet[151-176]{grasby2002}.} alongside lettering, as dating
criteria.\footnote{ For an extensive list of dating conventions cf. \citet[Chapter 20]{Manzella1987}. }
Panciera’s recent article, in particular, advocates the significance of the public context and visibility of
inscriptions.\footnote{ ``I would propose to regard as an ‘inscription’ any particular type of written human
communication of the sort that we would today call unidirectional…not being addressed to a person or to a group but to
a collectivity, and for this reason is made with the location, writing technique, graphic form and impagination, mode
and register of expression chosen because they are most suitable to the attainment of its intended goal.'' \citep[8]{panciera
2012}.} In this vein, it is worth considering how the physical characteristics of an inscription belong within a
broader assessment of a culture of writing. Is palaeography the study of letterforms alone? Was a focus on letterforms
a deliberate choice or a product of the traditional constraints in accessibility and publication of epigraphic
materials? The following case study from Aphrodisias will explore these questions further.

\subsection{Outline}

\noindent The focus of this paper is a series of rather unimpressive column dedications from the temple of Aphrodite at
Aphrodisias. While the information recorded in the texts is unremarkable, the journey of these inscriptions from
their publication as part of an epigraphic corpus to a recent digital publication reveals an epic transformation in the
format and approach to these materials. This study will begin with brief overview of previous published editions
(e.g. \citet{mama8}) followed by an assessment of the information available on the current Inscriptions of
Aphrodisias (2007) website. Through careful assessment of text (including formula, vocabulary and spelling) and its
presentation (the arrangement of the text, use of decoration and spaces, as well as lettering) in the images provided,
this survey will demonstrate how the availability of published images and inclusion of dating criteria have increased
the amount of information available whilst also adding clarity to the process of dating an inscription. By examining
how we define and use palaeography to evaluate inscriptions on the Inscriptions of Aphrodisias website \citep{ReynoldsRouecheBodard2007}, we
can observe how the discipline has evolved and what changes may be possible in approaches and the publication of
epigraphic material.

%%%%%%%%%
%IN THE DOC THERE IS A "BOX" WITH this text
%Key Questions:
%
%What elements of the text are available to us in different media?
%How do we record the dating process and criteria? 
%How do we represent (and attempt to reconcile) contradictory information?
%%%%%%%%%%%

\section{Publishing inscriptions: A Case Study of Column Dedications from The Temple Of Aphrodite at Aphrodisias}

\subsection{A brief introduction to the inscriptions IAph 1.4-1.6.}

\noindent Three inscriptions, each of which record the dedication of a column at the Temple of Aphrodite, will be the subject of
this survey. The first two texts were noted as early as the 18th century and copied in a notebook by the British
Architect Deering in the early 19th century. A third version of the text, was uncovered during excavations at the
temple site by the French Engineer Gaudin in the early 20th century.\footnote{ A comprehensive history of the history
and bibliography, as well as a description of the resources can be found on the IAph website:
\url{http://insaph.kcl.ac.uk/iaph2007/iAph010006.html}. } The inscriptions are recorded on tabella ansata (0.745 x 0.465m) as
part of fluted marble columns of the peristasis, some of these have been reconstructed in modern restorations on the
site.

These texts can be dated through a number of different criteria. Contextual association with construction of the
temple, dates between the end of the 1st century BC and the early 1st century AD, though the land was clearly in use
well before this time. \footnote{\citet[41-42]{smith1996} and \citet[37]{reynolds1990}. References to Sulla’s dedications are in
Appian BC 1.97. Caesar and Augustus’ acknowledgements of the sacred space are evident in IAph 8.27, 8.31 and 1.1. }
This is corroborated by coins depicting the temple, which date from 2 BC- 14 AD.\footnote{ Coin type 41 depicts
Augustus (OBV) and the temple of Aphrodite (REV) in MacDonald 1992, Plate V R131. } Prosopography is also informative:
Gaius Julius Zoilos (freed by Caesar or Octavian), who dedicated the theatre at Aphrodisias ca. 28 BC, also dedicated
the naos of the temple (IAph 1.2), perhaps posthumously. \footnote{ The titles on the architrave inscription, soter and
euergetes, are absent from Zoilos’ other inscriptions and may imply a posthumous dedication, perhaps by the boule and
demos \citep[38]{reynolds 1990}.} The benefactor of the columns in this case study, Eumachos Diogenes, also comes from an
established family that flourished well into the 2nd century AD.\footnote{18 The success of Eumachus Diogenes’ family,
which included the first known Aphrodisian to hold a procuratorship in the 2nd century AD, reveals an enduring
significance for his family monuments in the city \citep[327-334]{reynolds1999}.} Finally, the formula of the text the
inscription and the vocabulary, particularly the term philokaisar, suggest a late Republican or Augustan
date.\footnote{ The formula of the dedication, which lists the benefactor first is evident throughout the Hellenistic
period in Aphrodisias (and Asia Minor) until the early Imperial Period, after which a new formula (beginning with
recipients (e.g. Aphrodite and Imperial recipients) is predominant (cf. \citet[4-7]{graham2013}). The Augustan and early
Tiberian uses of the philokaisar are known from a dedication at Ioulis dated ca. 27 BC - AD 14 (SEG XLVIII (1998) no.
1129) and in a monument to Ti. Cl. Drusus in Patara (SEG XLIV (1994) no. 1205). For significance and date of the title
see \citet[101-105]{buraselis2001}.} Before encountering the inscriptions face to face, the dating, function and meaning of
these inscriptions appears to be quite straightforward. So let us examine the experience of viewing these
inscriptions various published formats.

\subsection{Publishing IAPH 1.4 and 1.5 in MAMA VIII (nos. 347 \& 348): A series of copies?}
\noindent Cormack published the first comprehensive catalogue of inscriptions from Aphrodisias as part of his Monumenta Asiae
Minoris Antiqua volume VIII in 1962. This was a great undertaking that included texts as well as images. Two of the
three texts (IAph 1.4 \& 1.5) were published here as MAMA 437 and MAMA 438 respectively, while the third text was
merely mentioned as a further copy.

\begin{quotation}
\noindent MAMA 437\\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηναγό-\\
ρου τοῦ Ἀθηναγόρου\\
τοῦ Εὐμάχου Διογένη-\\
ς φιλόκαισαρ καὶ Ἀμιὰς\\
Διονυσίου φύσι δὲ Ἀδρά<σ>του\\
τοῦ Μόλωνος Ὀλυνπιὰς\\
τὸν κίονα θεᾷ Ἀφροδίτῃ\\
καὶ τῷ δήμῳ.
\end{otherlanguage*}
\end{quotation}

\begin{quotation}
\noindent MAMA 438 \\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηνα-\\
γόρου τοῦ Ἀθηναγό-\\
ρου τοῦ Εὐμάχου Δι-\\
ογένης φιλόκαισαρ\\
καὶ Ἀμμιὰς Διονυσί-\\
ου φύσι δὲ Ἀδράστου\\
τοῦ Μόλωνος Ὀλυν-\\
πιὰς τὸν κίονα θεᾷ\\
Ἀφροδίτῃ καὶ τῷ\\
δήμῳ.
\end{otherlanguage*}
\end{quotation}

\begin{quotation}
\emph{Translation by author:}\\
Eumachos Diogenes, son of Athenagoras, the son of Athenagoras, the son of Eumachos, devoted to Caesar, and Ammias Olympias, daughter of Dionysius, the natural daughter of Adrastus, the son of Molon, (dedicated) a column to the goddess Aphrodite and the People.  
\end{quotation}

The two published texts reveal similar inscriptions with a few variations.\footnote{ MAMA 437 and 438 are represented in
their digital format as published the Packhum website \ (Aphrodisias 108 and 109 (respectively)).
\url{http://epigraphy.packhum.org/inscriptions/main?url=oi\%3Fikey\%3D256987}} IAph 1.4 (MAMA 437) has a slightly less
impressive ordinatio, particularly in line 4 with an odd line break, a misspelling of Ammias (as Amias) and Adrastus
(line 5) with a sigma omitted. It is difficult to determine how or why these errors occurred without indications of
spacing or an examination of the stone. \ One could argue that spelling and arrangement of the text were not important
to a broader audience at Aphrodisias, however, this is negated by a second copy of the dedication (IAph 1.5, MAMA 438),
which portrays a more skillful execution and arrangement of the text. Spelling errors are rectified and the arrangement
of the text creates a distinction in Line 5 between Eumachus and Ammias, and for the demos, which is isolated on the
last line. These subtle variations only emerge with a close reading of the text, and it would be easy to overlook them.
The published format of the texts also makes them appear similar in lettering and arrangement (e.g. left indentation).

@@

The third text version of this text is mentioned in the commentary and based on the information provided (two similar
texts and a reference to a ``copy'') one would expect that it was very similar. The same editorial practice is employed
in the parallel column dedications from a different benefactor (I.Aph 7 \& I.Aph 8), where 2nd version of the text is
noted (IAph 1.8) but only one (I.Aph 1.7, MAMA 450) was published. The apparent verisimilitude of the column
dedications is corroborated by the published images, which depict two inscriptions (MAMA 437 and MAMA 450) with similar
late 1st c. BC/early 1st c. AD lettering. When publishing a large corpus of inscriptions, referring to copies is
understandable. However, treating texts as ``copies'' as opposed to individual monuments can, and in this case will,
prove problematic.\footnote{ For another study of ``copy'' inscriptions at Aphrodisias see Graham 2015 (forthcoming).}
While MAMA VIII offered a broader audience access to the inscriptions at Aphrodisias, the approach to the inscriptions
and the quality of the images also imposed limitations that made it difficult to analyse the physical elements of the
inscription or to question the proposed texts and restorations.

\subsection{Living in a Digital World: The Publication of Column Dedications in IAph 2007 (IAph 1.4.)}
The Inscriptions of Aphrodisias project was a groundbreaking endeavour both for publishing epigraphic materials and
employing Epi Doc conventions with XML for marking up ancient documents. Its success has inspired a number of
significantly larger projects with international scope and collaboration. The approach to the materials is detailed
with categories covering physical characteristics, history of discovery and bibliography (which is important given the
wealth of superlative scholarship).\footnote{ It is worth noting that websites, however they develop, will remain
supplementary to published scholarship.} In addition to search functions, the page is more interactive, the texts
have been revisited, translations added, as well as commentary with weblinks to parallel texts. The organization of the
inscriptions by context allows the reader to view the inscriptions in context and to gain a better understanding of the
area’s ``epigraphic habit'' (Fig.~\ref{fig:1}).

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{PaperproposalforEAGLEfinal-img001.png}
\caption{Screenshot of IAph 4 webpage}
\label{fig:1}
\end{figure}

@@

The published texts of IAph 1.4 and MAMA 437 present one significant variation: the inclusion of space indicators within
the text. This practice allows the reader to see how the use of space relates with the text of the inscription,
particularly in the case of line 4, where challenges in the text (misspelling of Ammias and carrying over of a single
letter from the previous line), can be observed in the inclusion of a vacat at the end of the line.\footnote{ Vacats,
more sparingly employed in late Republican and Augustan inscriptions at Aphrodisias, usually serve a grammatical and/or
decorative function (e.g. giving distinction to a name or key elements/individuals in the text (cf. building
dedications at the Sebasteion IAph 9.1, 9.25, 9.112). This vacat serves little grammatical function, and is likely to
be a result of the arrangement of the text/ omission in Ammias’ name. } The indentation on the last line also reveals
a left orientation of the line which is more common in Late Republican/Augustan texts at Aphrodisias (e.g. IAph 1.1,
1.7, 1.8 and 1.38), as opposed to the ``justified'' approach (indentations on both sides of the text to accentuate a name
or word), which is more common in later Imperial inscriptions at Aphrodisias.

\begin{quotation}
\noindent MAMA 437\\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηναγό-\\
ρου τοῦ Ἀθηναγόρου\\
τοῦ Εὐμάχου Διογένη-\\
ς φιλόκαισαρ καὶ Ἀμιὰς\\
Διονυσίου φύσι δὲ Ἀδρά<σ>του\\
τοῦ Μόλωνος Ὀλυνπιὰς\\
τὸν κίονα θεᾷ Ἀφροδίτῃ\\
καὶ τῷ δήμῳ.
\end{otherlanguage*}

\end{quotation}
\begin{quotation}
\noindent IAph 1.4.\\
\begin{otherlanguage*}{greek}
\noindent Εὔμαχος Ἀθηναγό-\\
ρου τοῦ Ἀθηναγόρου\\
τοῦ Εὐμάχου Διογένη-\\
ς Φιλόκαισαρ καὶ Ἀμιὰς\end{otherlanguage*} v.\\
\begin{otherlanguage*}{greek}Διονυσίου φύσι δὲ Ἀδρά<σ>του\\
τοῦ Μόλωνος Ὀλυνπιὰς\\
τὸν κίονα θεᾷ Ἀφροδίτῃ\\
καὶ τῷ Δήμῳ
\end{otherlanguage*} vacat
\end{quotation}

High quality images on the website facilitate connections between the published text and the inscription. In the image
of IAph 1.4 (Fig.~\ref{fig:2}), one can to observe how descriptive elements manifest themselves in an
inscription: how varying letter sizes (2-3 cm) represent a lack of uniformity and crowding as space becomes more
cramped, and how, after the beautifully spaced upper lines (lines 1-3) the carver struggled (line 4) to fit the letters
in, possibly noticing his spelling error only when there was extra space at the end of the line. \
